---
title: "Global Carbon Estimates"
author: "Kaydee S. Barker"
date: '2023-09-22'
output: 
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
rm(list=ls()) #clear global environment/workspace
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, include=FALSE}
library(readr) #read files
library(tidyverse) #Tidy packages
library(dplyr)
library(sf) #Spatial package that can read and create shapefiles 
library(terra) #Spatial package that can work with raster files
library(ggplot2) #Pretty plots

```

## Ecosystem Cover Data

Biomes containing arbuscular mycorrhizal (AM) fungi include croplands, temperate grasslands, savannas and shrublands (i.e. tropical grasslands), temperate forests, and tropical forests. I used data from the United Nations Food and Agriculture Organization (FAO)'s Global Land Cover - SHARE database ([Latham et al. 2014](https://www.fao.org/uploads/media/glc-share-doc.pdf)) and proportions from [Dinerstein et al. 2017](https://doi.org/10.1093/biosci/bix014) to determine total area in square kilometers for each biome with AM fungi.

```{r cover}

# Download total land cover map from FAO
#download.file("https://storage.googleapis.com/fao-maps-catalog-data/uuid/ba4526fd-cdbf-4028-a1bd-5a559c4bff38/resources/GlcShare_v10_Dominant.zip", destfile= "FAO_covermap.zip")
#unzip("FAO_covermap.zip", exdir = ".")
#file.remove("FAO_covermap.zip")

biomes <- rast("glc_shv10_DOM.Tif")

plot(biomes) #look at map

#make df of frequency of each value in raster
landcover <- as.data.frame(freq(biomes)) %>%
  rename("Area" = "count")

landcover$Ecosystem <- c("Null", "Artificial", "Cropland", "Grassland", "Tree Covered", "Shrubland", "Herbaceous Wetland", "Mangrove", "Sparse Veg", "Bare Soil", "Snow & Glaciers", "Water")

#Split out and add ecosystems based on proportion

#Tropical forests - based on FAO 2020
tropfor <- data.frame(1,12,71015243*0.45, "Tropical Forest")
names(tropfor) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, tropfor)

#Temperate forests - based on FAO 2020
tempfor <- data.frame(1,12,71015243*0.16, "Temperate Forest")
names(tempfor) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, tempfor)

#Temperate grasslands - based on Dinerstein et al. 2017
tempgr <- data.frame(1,12,33129685*0.33, "Temperate Grassland")
names(tempgr) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, tempgr)

#Savannas - based on Dinerstein et al. 2017, plus shrub
savshrub <- data.frame(1,12,(33129685*0.67)+18337497, "Savanna & Shrubland")
names(savshrub) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, savshrub)

AMcover <- landcover[,c(3,4)] %>%
  filter(Ecosystem != "Null") %>% # filters out
  filter(Ecosystem != "Artificial") %>%
  filter(Ecosystem != "Grasslands") %>%
  filter(Ecosystem != "Tree Covered") %>%
  filter(Ecosystem != "Shrublands") %>%
  filter(Ecosystem != "Herbaceous Wetlands") %>%
  filter(Ecosystem != "Mangroves") %>%
  filter(Ecosystem != "Sparse Veg") %>%
  filter(Ecosystem != "Bare Soil") %>%
  filter(Ecosystem != "Snow & Glaciers") %>%
  filter(Ecosystem != "Water")

```

## Carbon Data

### Plant Carbon

Aboveground carbon (C) and root C were taken from previous C estimates and root:shoot ratios ([Robinson 2007](https://royalsocietypublishing.org/doi/10.1098/rspb.2007.1012), [Liu et al. 2015](https://www.nature.com/articles/nclimate2581)). All C data is in petagrams (Pg) of C.

```{r plant C data}

#aboveground and root C data from Robinson 2007, Liu et al. 2015
#used proportion of root:shoot from Robinson 2007 to fill in Liu et al. 2015 root C estimates - in Pg C
plantC <- read_csv("plantC.csv")

```

### AM Fungal Carbon

To estimate the contribution of AM fungal biomass to global C pools, I used the estimated amount of AM extra-radical hyphal biomass C in grams per square meter (g/m^2) from [Soudzilovskaia et al. 2015](https://onlinelibrary.wiley.com/doi/10.1111/nph.13447), scaled it to total biome area, and converted the units to Pg of C. Additionally, I included estimated C allocation from plants to mycorrhizae per year based on vegetation type averages from [Hawkins et al. 2023](https://doi.org/10.1016/j.cub.2023.02.027) and scaled to total biome area for a Pg of C per year estimate.


```{r AM C data}

#merge with cover, add ERH C based on Soudzilovskaia et al. 2015 - g/m-2 estimates. Different than C flux from Hawkins et al. 2023
coverC <- merge(AMcover, plantC, 
	by="Ecosystem", all.x = T) %>% #merge 
  mutate(AMbio_C = (Area*1000000)*0.0000000000000475) #add ERH

#C flux per veg type via Hawkins et al. 2023

#Croplands
cropAM <- coverC %>%
  filter(Ecosystem == "Cropland") %>%
  mutate(Allo_C = AGB_C * 0.033) 
coverC2 <- merge(coverC, cropAM)

#Savannas & shrublands
ssAM <- coverC %>%
  filter(Ecosystem == "Savanna & Shrubland") %>%
  mutate(Allo_C = (AGB_C/2 * 0.065) + (AGB_C/2 * 0.023))
coverC2 <- rbind(coverC2, ssAM)

#Temperate grasslands
tgAM <- coverC %>%
  filter(Ecosystem == "Temperate Grassland") %>%
  mutate(Allo_C = AGB_C * 0.065) 
coverC2 <- rbind(coverC2, tgAM)

#Temperate forests
tfAM <- coverC %>%
  filter(Ecosystem == "Temperate Forest") %>%
  mutate(Allo_C = AGB_C * 0.023) 
coverC2 <- rbind(coverC2, tfAM)

#Tropical forests
trfAM <- coverC %>%
  filter(Ecosystem == "Tropical Forest") %>%
  mutate(Allo_C = AGB_C * 0.023) 
coverC2 <- rbind(coverC2, trfAM)

```

### Soil Organic Matter Fractions

To estimate how much C is stored in particulate organic matter (POM) or mineral-associated organic matter (MAOM) fractions across biomes globally, I used data from [Georgiou et al. 2022](https://www.nature.com/articles/s41467-022-31540-9). The top 30cm was selected to compare topsoil, and was scaled up 

```{r SOM C data}

#MAOM and POM C data from Georgiou et al. 2022
SOMC <- read_csv("MOC_synthesis.csv") %>%
  rename("Ecosystem" = "Vegetation") %>%
  rename("MAOM_C" = "SiltClayC") %>%
  select(Ecosystem, Top.depth, Bottom.depth, Bulk.C, MAOM_C, POM_C)
SOMC[SOMC == "Grassland"] <- "Temperate Grassland"
SOMC[SOMC == "Savanna"] <- "Savanna & Shrubland"
SOMC[SOMC == "Shrubland"] <- "Savanna & Shrubland"

SOMC_top30 <- SOMC %>%
  filter(Bottom.depth <= 30)

#Scale to area of biome, using average bulk density & sampling depth
coverC_top30 <- merge(coverC2, SOMC_top30) %>%
  mutate(PgC = (1.33*(Bottom.depth)*10000000000)*
           (Bulk.C/1000000000000000000)*Area) %>%
  mutate(MAOM_PgC = (1.33*(Bottom.depth)*10000000000)*
           (MAOM_C/1000000000000000000)*Area) %>%
  mutate(POM_PgC = (1.33*(Bottom.depth)*10000000000)*
           (POM_C/1000000000000000000)*Area)

Ctop30means <- coverC_top30 %>%
  group_by(Ecosystem) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)



#write_csv(coverC3, "Cpools.csv")
#write_csv(Cpoolmeans, "Cpoolmeans.csv")

```


```{r}


```

