---
title: "Global Carbon Estimates"
author: "Kaydee S. Barker"
date: '2023-09-22'
output: 
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
rm(list=ls()) #clear global environment/workspace
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}

library(readr) #read files
library(tidyverse) #Tidy packages
library(dplyr)
library(sf) #Spatial package that can read and create shapefiles 
library(ggplot2) #Pretty plots
library(devtools) #install packages
library(RColorBrewer)
library(ggthemes)

```

## Ecosystem Cover Data

Biomes containing arbuscular mycorrhizal (AM) fungi include croplands, temperate grasslands, savannas and shrublands (i.e. tropical grasslands), temperate forests, and tropical forests. I used data from the United Nations Food and Agriculture Organization (FAO)'s Global Land Cover - SHARE database ([Latham et al. 2014](https://www.fao.org/uploads/media/glc-share-doc.pdf)) and proportions from [Dinerstein et al. 2017](https://doi.org/10.1093/biosci/bix014) to determine total area in square kilometers for each biome, Then used data from [Steidinger et al. 2019](https://doi.org/10.1038/s41586-019-1411-2) to determine the proportion of AM vs. ectomycorrhizal (EcM) fungal trees.

```{r cover}

#Download total land cover map from FAO if not downloaded already
if(file.exists('/glc_shv10_DOM.Tif')){
  print('downloaded')
  }else{
  download.file("https://storage.googleapis.com/fao-maps-catalog-data/uuid/ba4526fd-cdbf-4028-a1bd-5a559c4bff38/resources/GlcShare_v10_Dominant.zip", destfile= "FAO_covermap.zip")
unzip("FAO_covermap.zip", exdir = ".")
file.remove("FAO_covermap.zip")}

biomes <- rast("glc_shv10_DOM.Tif")

#Make df of frequency of each value in raster
landcover <- as.data.frame(freq(biomes)) %>%
  rename("Area" = "count") 

landcover$Ecosystem <- c("Null", "Artificial", "Cropland", "Grassland", "Tree Covered", "Shrubland", "Herbaceous Wetland", "Mangrove", "Sparse Veg", "Bare Soil", "Snow & Glaciers", "Water")

#Download total land cover map from EcoRegions2017 if not downloaded already
if(file.exists('/Ecoregions2017.shp')){
  print('downloaded')
  }else{
  download.file("https://storage.googleapis.com/teow2016/Ecoregions2017.zip", 
                destfile= "Ecoregions2017.zip", method= "libcurl", 
                quiet = FALSE, mode = "w", cacheOK = TRUE,
                options(timeout = max(300, getOption("timeout"))))
unzip("Ecoregions2017.zip", exdir = ".")
file.remove("Ecoregions2017.zip")}

ecoregions <- st_read("Ecoregions2017.shp")

library(foreign)
ecoregions_df <- read.dbf("Ecoregions2017.dbf")

#Calculate area of each ecoregion into sq km
st_crs(ecoregions) #check coordinate system
ecoregions <- st_make_valid(ecoregions) #fix invalid geometries

#Transform to a projected CRS (e.g., UTM) for accurate area calculations
#Replace EPSG:4326 with planar CRS (e.g., EPSG:3395 for World Mercator)
projected_ecoregions <- st_transform(ecoregions, crs = 3395) # EPSG:3395 for World Mercator

#Calculate area in square meters and convert to square kilometers
projected_ecoregions$area_sq_m <- st_area(projected_ecoregions)
projected_ecoregions$area_sq_km <- as.numeric(projected_ecoregions$area_sq_m) / 1e6


#Split out and add ecosystems based on proportion - data from FAO 2020, Dinerstein et al. 2017, Steidinger et al. 2019 for AM vs. EcM ratio in forests

#Proportions data
Dinerstein <- data.frame(
  Biome = c("1. Tropical and subtropical moist broadleaf forests", "2. Tropical and subtropical dry broadleaf forests", "3. Tropical and subtropical coniferous forests", "4. Temperate broadleaf and mixed forests", "5. Temperate conifer forests", "6. Boreal forests or taiga", "14. Mangroves", "Forested subtotal", "7. Tropical grasslands, savannas, and shrublands", "8. Temperate grasslands, savannas, and shrublands", "9. Flooded grasslands and savannas", "10. Montane grasslands and shrublands", "11. Tundra", "12. Mediterranean forests, woodlands, and scrub", "13. Desert and xeric shrublands", "Nonforested subtotal"),
  Percent = c(14.4, 2.9, 0.5, 9.3, 2.8, 11.4, 0.2, 41.3, 15.8, 7.8, 0.9, 3.6, 8.7, 2.4, 19.3, 58.5)
)
Steidinger <- read.csv('AM_EM_proportions.csv') #read in data

#Tropical and subtropical forests
tropfor <- data.frame(1,12,
            landcover$Area[landcover$Ecosystem == "Tree Covered"]*
              ((Dinerstein$Percent[Dinerstein$Biome == "1. Tropical and subtropical moist broadleaf forests"]*(Steidinger$median[Steidinger$group_id == "Wet_trop_broadleaf_AM"])/100) +
                 (Dinerstein$Percent[Dinerstein$Biome == "2. Tropical and subtropical dry broadleaf forests"]*(Steidinger$median[Steidinger$group_id == "Dry_trop_broadleaf_AM"])/100) +
              (Dinerstein$Percent[Dinerstein$Biome == "3. Tropical and subtropical coniferous forests"]*(Steidinger$median[Steidinger$group_id == "Trop_conifer_AM"])/100)) /
                 Dinerstein$Percent[Dinerstein$Biome == "Forested subtotal"],
            "Tropical Forest")
names(tropfor) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, tropfor)

#Temperate forests
tempfor <- data.frame(1,13,landcover$Area[landcover$Ecosystem == "Tree Covered"]*
          ((Dinerstein$Percent[Dinerstein$Biome == "4. Temperate broadleaf and mixed forests"]*(Steidinger$median[Steidinger$group_id == "Temp_mixed_AM"])/100) +
            (Dinerstein$Percent[Dinerstein$Biome == "5. Temperate conifer forests"]*(Steidinger$median[Steidinger$group_id == "Temp_conifer_AM"])/100)) /
            Dinerstein$Percent[Dinerstein$Biome == "Forested subtotal"],
                      "Temperate Forest")
names(tempfor) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, tempfor)

#Temperate grasslands
tempgr <- data.frame(1,14,
                     (landcover$Area[landcover$Ecosystem == "Grassland"] + 
                        landcover$Area[landcover$Ecosystem == "Shrubland"]) *
                      ((Dinerstein$Percent[Dinerstein$Biome == "8. Temperate grasslands, savannas, and shrublands"]*((Steidinger$median[Steidinger$group_id == "Temp_savannah_AM"] + Steidinger$median[Steidinger$group_id == "Xeric_shrub_AM"] + 85)/3)/100)) /
                        Dinerstein$Percent[Dinerstein$Biome == "Nonforested subtotal"],
                     "Temperate Grassland, Shrubland, and Savanna")
names(tempgr) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, tempgr)

#Savannas - based on Dinerstein et al. 2017, plus shrub
savshrub <- data.frame(1,15,
                       (33129685*0.67)+18337497, "Savanna & Shrubland")
names(savshrub) <- c("layer","value","Area","Ecosystem")
landcover <- rbind(landcover, savshrub)

AMcover <- landcover[,c(3,4)] %>%
  filter(Ecosystem != "Null") %>% # filters out
  filter(Ecosystem != "Artificial") %>%
  filter(Ecosystem != "Grasslands") %>%
  filter(Ecosystem != "Tree Covered") %>%
  filter(Ecosystem != "Shrublands") %>%
  filter(Ecosystem != "Herbaceous Wetlands") %>%
  filter(Ecosystem != "Mangroves") %>%
  filter(Ecosystem != "Sparse Veg") %>%
  filter(Ecosystem != "Bare Soil") %>%
  filter(Ecosystem != "Snow & Glaciers") %>%
  filter(Ecosystem != "Water")

```

## Carbon Data

### Soil Data

The [International Soil Carbon Network](http://iscn.fluxdata.org/) has made a global dataset of soil characteristics available for public use ([Nave et al. 2016](https://www.osti.gov/biblio/1305039)).

```{r soil data, error=FALSE}

#Github package for ISCN data - no longer supported
#install_github("ISCN/SOCDRaHR2")
#library(SOCDRaH2)
#ISCN <- ISCN3() #download ISCN data

#if/else loop for file download if the file isn't already downloaded
if(file.exists('/ISCN.csv')){ #if a file exists in working directory
		print('downloaded') #print result
	} else{ #if it doesn't exist
		download.file("https://pasta.lternet.edu/package/data/eml/edi/1160/1/4af719a84f8981fcc63f1f92760cb253", 
		              destfile= "ISCN.csv") #download file from link and name it
	} #end loop

ISCN <- read.csv("ISCN.csv", header=F 
          ,skip=1
            ,sep=";"  
                ,quot='"' 
        , col.names=c(
                    "ISCN.1.hyphen.1..paren.2015.hyphen.12.hyphen.10.paren.",     
                    "dataset_name_sub",     
                    "dataset_name_soc",     
                    "lat..paren.dec..deg.paren.",     
                    "long..paren.dec..deg.paren.",     
                    "datum..paren.datum.paren.",     
                    "state..paren.state_province.paren.",     
                    "country..paren.country.paren.",     
                    "site_name",     
                    "observation_date..paren.YYYY.hyphen.MM.hyphen.DD.paren.",     
                    "profile_name",     
                    "layer_name",     
                    "layer_top..paren.cm.paren.",     
                    "layer_bot..paren.cm.paren.",     
                    "hzn_desgn_other",     
                    "hzn",     
                    "hzn_desgn",     
                    "layer_note",     
                    "color",     
                    "vegclass_local",     
                    "soil_taxon",     
                    "soil_series",     
                    "bd_method",     
                    "bd_samp..paren.g.cm.hyphen.3.paren.",     
                    "bd_tot..paren.g.cm.hyphen.3.paren.",     
                    "bd_whole..paren.g.cm.hyphen.3.paren.",     
                    "bd_other..paren.g.cm.hyphen.3.paren.",     
                    "bdNRCS_prep_code",     
                    "cNRCS_prep_code",     
                    "c_method",     
                    "c_tot..paren.percent.paren.",     
                    "oc..paren.percent.paren.",     
                    "loi..paren.percent.paren.",     
                    "n_tot..paren.percent.paren.",     
                    "c_to_n..paren.mass.ratio.paren.",     
                    "soc..paren.g.cm.hyphen.2.paren.",     
                    "soc_carbon_flag",     
                    "soc_method",     
                    "ph_method",     
                    "ph_cacl",     
                    "ph_h2o",     
                    "ph_other",     
                    "caco3..paren.percent.paren.",     
                    "sand_tot_psa..paren.percent.paren.",     
                    "silt_tot_psa..paren.percent.paren.",     
                    "clay_tot_psa..paren.percent.paren.",     
                    "wpg2_method",     
                    "wpg2..paren.percent.paren.",     
                    "cat_exch..paren.cmol.H.plus..kg.hyphen.1.paren.",     
                    "al_dith..paren.specified.by.al_fe_units.paren.",     
                    "al_ox..paren.specified.by.al_fe_units.paren.",     
                    "al_other..paren.specified.by.al_fe_units.paren.",     
                    "fe_dith..paren.specified.by.al_fe_units.paren.",     
                    "fe_ox..paren.specified.by.al_fe_units.paren.",     
                    "fe_other..paren.specified.by.al_fe_units.paren.",     
                    "mn_dith..paren.specified.by.al_fe_units.paren.",     
                    "mn_ox..paren.specified.by.al_fe_units.paren.",     
                    "mn_other..paren.specified.by.al_fe_units.paren.",     
                    "al_fe_units..paren.extract_units.paren.",     
                    "al_fe_method",     
                    "ca_al..paren.specified.by.bc_units.paren.",     
                    "ca_ext..paren.specified.by.bc_units.paren.",     
                    "k_ext..paren.specified.by.bc_units.paren.",     
                    "mg_ext..paren.specified.by.bc_units.paren.",     
                    "na_ext..paren.specified.by.bc_units.paren.",     
                    "bc_units..paren.extract_units.paren.",     
                    "bc_method",     
                    "base_sum..paren.specified.by.cec_h_units.paren.",     
                    "cec_sum..paren.specified.by.cec_h_units.paren.",     
                    "ecec..paren.specified.by.cec_h_units.paren.",     
                    "cec_h_units..paren.extract_units.paren.",     
                    "bs..paren.percent.paren.",     
                    "bs_sum..paren.percent.paren.",     
                    "h_ext..paren.specified.by.metal_ext_units.paren.",     
                    "zn_ext..paren.specified.by.metal_ext_units.paren.",     
                    "metal_ext_units..paren.extract_units.paren.",     
                    "metal_ext_method",     
                    "p_bray..paren.specified.by.p_units.paren.",     
                    "p_ox..paren.specified.by.p_units.paren.",     
                    "p_meh..paren.specified.by.p_units.paren.",     
                    "p_other..paren.specified.by.p_units.paren.",     
                    "p_units..paren.extract_units.paren.",     
                    "p_method",     
                    "root_quant_size",     
                    "root_weight..paren.g.paren.",     
                    "v_15n..paren.‰.paren.",     
                    "v_13c..paren.‰.paren.",     
                    "v_14c..paren.‰.paren.",     
                    "v_14c_sigma..paren.‰.paren.",     
                    "v_14c_age..paren.BP.paren.",     
                    "v_14c_age_sigma..paren.BP.paren.",     
                    "fraction_modern",     
                    "fraction_modern_sigma",     
                    "textureClass",     
                    "locator_parent_alias"    ), check.names=TRUE)


#fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
if (class(ISCN$lat..paren.dec..deg.paren.)!="factor") ISCN$lat..paren.dec..deg.paren.<- as.factor(ISCN$lat..paren.dec..deg.paren.)
if (class(ISCN$long..paren.dec..deg.paren.)!="factor") ISCN$long..paren.dec..deg.paren.<- as.factor(ISCN$long..paren.dec..deg.paren.)
if (class(ISCN$datum..paren.datum.paren.)!="factor") ISCN$datum..paren.datum.paren.<- as.factor(ISCN$datum..paren.datum.paren.)
if (class(ISCN$layer_top..paren.cm.paren.)!="factor") ISCN$layer_top..paren.cm.paren.<- as.factor(ISCN$layer_top..paren.cm.paren.)
if (class(ISCN$layer_bot..paren.cm.paren.)!="factor") ISCN$layer_bot..paren.cm.paren.<- as.factor(ISCN$layer_bot..paren.cm.paren.)
if (class(ISCN$vegclass_local)!="factor") ISCN$vegclass_local<- as.factor(ISCN$vegclass_local)
if (class(ISCN$soil_taxon)!="factor") ISCN$soil_taxon<- as.factor(ISCN$soil_taxon)
if (class(ISCN$soil_series)!="factor") ISCN$soil_series<- as.factor(ISCN$soil_series)
if (class(ISCN$bd_method)!="factor") ISCN$bd_method<- as.factor(ISCN$bd_method)
if (class(ISCN$bd_samp..paren.g.cm.hyphen.3.paren.)!="factor") ISCN$bd_samp..paren.g.cm.hyphen.3.paren.<- as.factor(ISCN$bd_samp..paren.g.cm.hyphen.3.paren.)
if (class(ISCN$bd_tot..paren.g.cm.hyphen.3.paren.)!="factor") ISCN$bd_tot..paren.g.cm.hyphen.3.paren.<- as.factor(ISCN$bd_tot..paren.g.cm.hyphen.3.paren.)
if (class(ISCN$bd_whole..paren.g.cm.hyphen.3.paren.)!="factor") ISCN$bd_whole..paren.g.cm.hyphen.3.paren.<- as.factor(ISCN$bd_whole..paren.g.cm.hyphen.3.paren.)
if (class(ISCN$bd_other..paren.g.cm.hyphen.3.paren.)!="factor") ISCN$bd_other..paren.g.cm.hyphen.3.paren.<- as.factor(ISCN$bd_other..paren.g.cm.hyphen.3.paren.)
if (class(ISCN$c_method)!="factor") ISCN$c_method<- as.factor(ISCN$c_method)
if (class(ISCN$c_tot..paren.percent.paren.)!="factor") ISCN$c_tot..paren.percent.paren.<- as.factor(ISCN$c_tot..paren.percent.paren.)
if (class(ISCN$oc..paren.percent.paren.)!="factor") ISCN$oc..paren.percent.paren.<- as.factor(ISCN$oc..paren.percent.paren.)
if (class(ISCN$loi..paren.percent.paren.)!="factor") ISCN$loi..paren.percent.paren.<- as.factor(ISCN$loi..paren.percent.paren.)
if (class(ISCN$n_tot..paren.percent.paren.)!="factor") ISCN$n_tot..paren.percent.paren.<- as.factor(ISCN$n_tot..paren.percent.paren.)
if (class(ISCN$c_to_n..paren.mass.ratio.paren.)!="factor") ISCN$c_to_n..paren.mass.ratio.paren.<- as.factor(ISCN$c_to_n..paren.mass.ratio.paren.)
if (class(ISCN$soc..paren.g.cm.hyphen.2.paren.)!="factor") ISCN$soc..paren.g.cm.hyphen.2.paren.<- as.factor(ISCN$soc..paren.g.cm.hyphen.2.paren.)
if (class(ISCN$soc_carbon_flag)!="factor") ISCN$soc_carbon_flag<- as.factor(ISCN$soc_carbon_flag)
if (class(ISCN$soc_method)!="factor") ISCN$soc_method<- as.factor(ISCN$soc_method)
if (class(ISCN$ph_method)!="factor") ISCN$ph_method<- as.factor(ISCN$ph_method)
if (class(ISCN$ph_cacl)!="factor") ISCN$ph_cacl<- as.factor(ISCN$ph_cacl)
if (class(ISCN$ph_h2o)!="factor") ISCN$ph_h2o<- as.factor(ISCN$ph_h2o)
if (class(ISCN$ph_other)!="factor") ISCN$ph_other<- as.factor(ISCN$ph_other)
if (class(ISCN$sand_tot_psa..paren.percent.paren.)!="factor") ISCN$sand_tot_psa..paren.percent.paren.<- as.factor(ISCN$sand_tot_psa..paren.percent.paren.)
if (class(ISCN$silt_tot_psa..paren.percent.paren.)!="factor") ISCN$silt_tot_psa..paren.percent.paren.<- as.factor(ISCN$silt_tot_psa..paren.percent.paren.)
if (class(ISCN$clay_tot_psa..paren.percent.paren.)!="factor") ISCN$clay_tot_psa..paren.percent.paren.<- as.factor(ISCN$clay_tot_psa..paren.percent.paren.)
if (class(ISCN$root_quant_size)!="factor") ISCN$root_quant_size<- as.factor(ISCN$root_quant_size)
if (class(ISCN$root_weight..paren.g.paren.)!="factor") ISCN$root_weight..paren.g.paren.<- as.factor(ISCN$root_weight..paren.g.paren.)
if (class(ISCN$textureClass)!="factor") ISCN$textureClass<- as.factor(ISCN$textureClass)

#data file 2
#if/else loop for file download if the file isn't already downloaded
if(file.exists('/ISCN_2.csv')){ #if a file exists in working directory
		print('downloaded') #print result
	} else{ #if it doesn't exist
		download.file("https://pasta.lternet.edu/package/data/eml/edi/1160/1/40527580cc045d33d9a5aaf728bf204e", 
		              destfile= "ISCN_2.csv") #download file from link and name it
	} #end loop

ISCN_2 <- read.csv("ISCN_2.csv", header=F 
          ,skip=1
            ,sep=";"  
                ,quot='"' 
        , col.names=c(
                    "ISCN.1.hyphen.1..paren.2015.hyphen.12.hyphen.10.paren.",     
                    "dataset_name_sub",     
                    "dataset_name_soc",     
                    "lat..paren.dec..deg.paren.",     
                    "long..paren.dec..deg.paren.",     
                    "datum..paren.datum.paren.",     
                    "state..paren.state_province.paren.",     
                    "country..paren.country.paren.",     
                    "observation_date..paren.YYYY.hyphen.MM.hyphen.DD.paren.",     
                    "site_name",     
                    "profile_name",     
                    "profile_zero_ref..paren.profile_zero_ref.paren.",     
                    "layer_top..paren.cm.paren.",     
                    "layer_bot..paren.cm.paren.",     
                    "total_lcount",     
                    "carbon_lcount",     
                    "soc_lcount",     
                    "soc_depth..paren.cm.paren.",     
                    "soc..paren.g.cm.hyphen.2.paren.",     
                    "soc_method",     
                    "soc_spatial_flag",     
                    "soc_carbon_flag",     
                    "soil_taxon",     
                    "soil_series",     
                    "elevation..paren.m.paren.",     
                    "ecoregion",     
                    "landuse..paren.landsat.paren.",     
                    "vegclass_local",     
                    "surface_veg",     
                    "landscape..paren.landscape.paren.",     
                    "landform..paren.landform.paren.",     
                    "v_2d_position..paren.2d_position.paren.",     
                    "landform_note",     
                    "aspect_deg..paren.degree.paren.",     
                    "slope..paren.percent.paren.",     
                    "drainagecl..paren.drainage.paren.",     
                    "map..paren.mm.paren.",     
                    "mat..paren.°C.paren.",     
                    "runoff..paren.runoff.paren.",     
                    "site_perm..paren.site_perm.paren.",     
                    "site_note",     
                    "thaw_depth_profile..paren.cm.paren.",     
                    "locator_alias",     
                    "locator_parent_alias"    ), check.names=TRUE)

#fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
if (class(ISCN_2$lat..paren.dec..deg.paren.)!="factor") ISCN_2$lat..paren.dec..deg.paren.<- as.factor(ISCN_2$lat..paren.dec..deg.paren.)
if (class(ISCN_2$long..paren.dec..deg.paren.)!="factor") ISCN_2$long..paren.dec..deg.paren.<- as.factor(ISCN_2$long..paren.dec..deg.paren.)
if (class(ISCN_2$datum..paren.datum.paren.)!="factor") ISCN_2$datum..paren.datum.paren.<- as.factor(ISCN_2$datum..paren.datum.paren.)
if (class(ISCN_2$layer_top..paren.cm.paren.)!="factor") ISCN_2$layer_top..paren.cm.paren.<- as.factor(ISCN_2$layer_top..paren.cm.paren.)
if (class(ISCN_2$layer_bot..paren.cm.paren.)!="factor") ISCN_2$layer_bot..paren.cm.paren.<- as.factor(ISCN_2$layer_bot..paren.cm.paren.)
if (class(ISCN_2$total_lcount)!="factor") ISCN_2$total_lcount<- as.factor(ISCN_2$total_lcount)
if (class(ISCN_2$carbon_lcount)!="factor") ISCN_2$carbon_lcount<- as.factor(ISCN_2$carbon_lcount)
if (class(ISCN_2$soc_lcount)!="factor") ISCN_2$soc_lcount<- as.factor(ISCN_2$soc_lcount)
if (class(ISCN_2$soc_depth..paren.cm.paren.)!="factor") ISCN_2$soc_depth..paren.cm.paren.<- as.factor(ISCN_2$soc_depth..paren.cm.paren.)
if (class(ISCN_2$soc..paren.g.cm.hyphen.2.paren.)!="factor") ISCN_2$soc..paren.g.cm.hyphen.2.paren.<- as.factor(ISCN_2$soc..paren.g.cm.hyphen.2.paren.)
if (class(ISCN_2$soc_method)!="factor") ISCN_2$soc_method<- as.factor(ISCN_2$soc_method)
if (class(ISCN_2$soc_spatial_flag)!="factor") ISCN_2$soc_spatial_flag<- as.factor(ISCN_2$soc_spatial_flag)
if (class(ISCN_2$soc_carbon_flag)!="factor") ISCN_2$soc_carbon_flag<- as.factor(ISCN_2$soc_carbon_flag)
if (class(ISCN_2$soil_taxon)!="factor") ISCN_2$soil_taxon<- as.factor(ISCN_2$soil_taxon)
if (class(ISCN_2$soil_series)!="factor") ISCN_2$soil_series<- as.factor(ISCN_2$soil_series)
if (class(ISCN_2$elevation..paren.m.paren.)!="factor") ISCN_2$elevation..paren.m.paren.<- as.factor(ISCN_2$elevation..paren.m.paren.)
if (class(ISCN_2$ecoregion)!="factor") ISCN_2$ecoregion<- as.factor(ISCN_2$ecoregion)
if (class(ISCN_2$landuse..paren.landsat.paren.)!="factor") ISCN_2$landuse..paren.landsat.paren.<- as.factor(ISCN_2$landuse..paren.landsat.paren.)
if (class(ISCN_2$vegclass_local)!="factor") ISCN_2$vegclass_local<- as.factor(ISCN_2$vegclass_local)
if (class(ISCN_2$surface_veg)!="factor") ISCN_2$surface_veg<- as.factor(ISCN_2$surface_veg)
if (class(ISCN_2$landscape..paren.landscape.paren.)!="factor") ISCN_2$landscape..paren.landscape.paren.<- as.factor(ISCN_2$landscape..paren.landscape.paren.)
if (class(ISCN_2$landform..paren.landform.paren.)!="factor") ISCN_2$landform..paren.landform.paren.<- as.factor(ISCN_2$landform..paren.landform.paren.)
if (class(ISCN_2$map..paren.mm.paren.)!="factor") ISCN_2$map..paren.mm.paren.<- as.factor(ISCN_2$map..paren.mm.paren.)
#if (class(ISCN_2$mat..paren.°C.paren.)!="factor") ISCN_2$mat..paren.°C.paren.<- as.factor(ISCN_2$mat..paren.°C.paren.)
if (class(ISCN_2$locator_alias)!="factor") ISCN_2$locator_alias<- as.factor(ISCN_2$locator_alias)
if (class(ISCN_2$locator_parent_alias)!="factor") ISCN_2$locator_parent_alias<- as.factor(ISCN_2$locator_parent_alias)

#data file 3
#if/else loop for file download if the file isn't already downloaded
if(file.exists('/ISCN_3.csv')){ #if a file exists in working directory
		print('downloaded') #print result
	} else{ #if it doesn't exist
		download.file("https://pasta.lternet.edu/package/data/eml/edi/1160/1/320e31ca911f187550ca2143c31fd408", 
		              destfile= "ISCN_3.csv") #download file from link and name it
	} #end loop

ISCN_3 <-read.csv("ISCN_3.csv", header=F 
          ,skip=1
            ,sep=";"  
                ,quot='"' 
        , col.names=c(
                    "ISCN.1.hyphen.1..paren.2015.hyphen.12.hyphen.10.paren.",     
                    "dataset_name",     
                    "dataset_type..paren.dataset_type.paren.",     
                    "curator_name",     
                    "curator_organization",     
                    "curator_email",     
                    "reference",     
                    "citation",     
                    "citation_usage",     
                    "acknowledgement",     
                    "acknowledgement_usage",     
                    "modification_date..paren.YYYY.hyphen.MM.hyphen.DD.paren."    ), check.names=TRUE)
		    

#data file 4
#if/else loop for file download if the file isn't already downloaded
if(file.exists('/ISCN_4.csv')){ #if a file exists in working directory
		print('downloaded') #print result
	} else{ #if it doesn't exist
		download.file("https://pasta.lternet.edu/package/data/eml/edi/1160/1/cdd0c7a4cac3f28d6d788c91f506775f", 
		              destfile= "ISCN_4.csv") #download file from link and name it
	} #end loop

ISCN_4 <-read.csv("ISCN_4.csv", header=F 
          ,skip=1
            ,sep=";"  
                ,quot='"' 
        , col.names=c(
                    "ISCN.1.hyphen.1..paren.2015.hyphen.12.hyphen.10.paren.",     
                    "dataset_name",     
                    "dataset_type..paren.dataset_type.paren.",     
                    "curator_name",     
                    "curator_organization",     
                    "curator_email",     
                    "contact_name",     
                    "contact_email",     
                    "dataset_description",     
                    "total_scount",     
                    "total_pcount",     
                    "total_lcount",     
                    "soc_scount",     
                    "soc_pcount",     
                    "soc_lcount",     
                    "soc_scount_ISCN",     
                    "soc_pcount_ISCN",     
                    "soc_lcount_ISCN",     
                    "modification_date..paren.YYYY.hyphen.MM.hyphen.DD.paren."    ), check.names=TRUE)

#look at data structure
str(ISCN) #structure


#select only usable data
soil_df <- ISCN[, c(3:6, 20, 24, 26, 32, 36, 37, 41, 44:46)] %>% #dataset name, lat/long, vegclass, BDs, SOC, pH, sand/silt/clay
  filter(bd_whole..paren.g.cm.hyphen.3.paren. != "NA") %>% #filter out NA values
  mutate(BD = case_when(bd_samp..paren.g.cm.hyphen.3.paren. != "NA" ~ bd_samp..paren.g.cm.hyphen.3.paren., #use bd_samp if it's not 0
            TRUE ~ bd_whole..paren.g.cm.hyphen.3.paren.)) #otherwise use bd_whole
  mutate(across(c(8,9,11:13), as.numeric)) #convert columns to numeric


```

### Plant Carbon

Aboveground carbon (C) and root C were taken from previous C estimates and root:shoot ratios ([Robinson 2007](https://royalsocietypublishing.org/doi/10.1098/rspb.2007.1012), [Liu et al. 2015](https://www.nature.com/articles/nclimate2581)). All C data is in petagrams (Pg) of C.

```{r plant C data}

#aboveground and root C data from Robinson 2007, Liu et al. 2015
#used proportion of root:shoot from Robinson 2007 to fill in Liu et al. 2015 root C estimates - in Pg C
plantC <- read_csv("plantC.csv")

```

### AM Fungal Carbon

To estimate the contribution of AM fungal biomass to global C pools, I used the estimated amount of AM extra-radical hyphal biomass C in grams per square meter (g/m^2) from [Soudzilovskaia et al. 2015](https://onlinelibrary.wiley.com/doi/10.1111/nph.13447), scaled it to total biome area, and converted the units to Pg of C. Additionally, I included estimated C allocation from plants to mycorrhizae per year based on vegetation type averages from [Hawkins et al. 2023](https://doi.org/10.1016/j.cub.2023.02.027) and scaled to total biome area for a Pg of C per year estimate.


```{r AM C data}

#merge with cover, add ERH C based on Soudzilovskaia et al. 2015 - g/m-2 estimates. Different than C flux from Hawkins et al. 2023
coverC <- merge(AMcover, plantC, 
	by="Ecosystem", all.x = T) %>% #merge 
  mutate(AMbio_C = (Area*1000000)*0.0000000000000475) #add ERH

#C flux per veg type via Hawkins et al. 2023

#Croplands
cropAM <- coverC %>%
  filter(Ecosystem == "Cropland") %>%
  mutate(Allo_C = AGB_C * 0.033) 
coverC2 <- merge(coverC, cropAM)

#Savannas & shrublands
ssAM <- coverC %>%
  filter(Ecosystem == "Savanna & Shrubland") %>%
  mutate(Allo_C = (AGB_C/2 * 0.065) + (AGB_C/2 * 0.023))
coverC2 <- rbind(coverC2, ssAM)

#Temperate grasslands
tgAM <- coverC %>%
  filter(Ecosystem == "Temperate Grassland") %>%
  mutate(Allo_C = AGB_C * 0.065) 
coverC2 <- rbind(coverC2, tgAM)

#Temperate forests
tfAM <- coverC %>%
  filter(Ecosystem == "Temperate Forest") %>%
  mutate(Allo_C = AGB_C * 0.023) 
coverC2 <- rbind(coverC2, tfAM)

#Tropical forests
trfAM <- coverC %>%
  filter(Ecosystem == "Tropical Forest") %>%
  mutate(Allo_C = AGB_C * 0.023) 
coverC2 <- rbind(coverC2, trfAM)

```

### Soil Organic Matter Fractions

To estimate how much C is stored in particulate organic matter (POM) or mineral-associated organic matter (MAOM) fractions across biomes globally, I used data from [Georgiou et al. 2022](https://www.nature.com/articles/s41467-022-31540-9). The top 30cm was selected to compare topsoil, and was scaled up to the area of global cover.

```{r SOM C data}

#MAOM and POM C data from Georgiou et al. 2022
SOMC <- read_csv("MOC_synthesis.csv") %>%
  rename("Ecosystem" = "Vegetation") %>%
  rename("MAOM_C" = "SiltClayC") %>%
  select(Ecosystem, Top.depth, Bottom.depth, Bulk.C, MAOM_C, POM_C)
SOMC[SOMC == "Grassland"] <- "Temperate Grassland"
SOMC[SOMC == "Savanna"] <- "Savanna & Shrubland"
SOMC[SOMC == "Shrubland"] <- "Savanna & Shrubland"

SOMC_top30 <- SOMC %>%
  filter(Bottom.depth <= 30)

#Scale to area of biome, using average bulk density & sampling depth
coverC_top30 <- merge(coverC2, SOMC_top30) %>%
  mutate(PgC = (1.33*(Bottom.depth)*10000000000)*
           (Bulk.C/1000000000000000000)*Area) %>%
  mutate(MAOM_PgC = (1.33*(Bottom.depth)*10000000000)*
           (MAOM_C/1000000000000000000)*Area) %>%
  mutate(POM_PgC = (1.33*(Bottom.depth)*10000000000)*
           (POM_C/1000000000000000000)*Area)

Ctop30means <- coverC_top30 %>%
  group_by(Ecosystem) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)



write_csv(coverC_top30, "Cpools30cm.csv")
write_csv(Ctop30means, "Cpoolmeans30cm.csv")

install.packages("gt")
library(gt)

C_table <- Ctop30means[, c(1:6,10:12)]

C_table <- C_table %>%
  column_to_rownames(var = "Ecosystem") %>%
  signif(., 4) #round to 4 significant figures

gt(C_table, 
   rownames_to_stub=TRUE)

```

## Plots

```{r plot AM bio}

#bar chart AM biomass C
ggplot(coverC_top30, aes(x=Ecosystem, y=AMbio_C, fill=Ecosystem)) + 
    geom_bar(position="dodge", stat = "identity", width = 0.7, colour="black") + #add proportion bars with black outline
    scale_x_discrete(guide = guide_axis(angle = 45)) +
    scale_fill_manual(values = wes_palette("Zissou1")) + #call from wesanderson library
    theme_few() #remove grid, white background, box around plot - from ggthemes

```


```{r plot AM allocation}

#bar chart AM allocation C
ggplot(coverC_top30, aes(x=Ecosystem, y=Allo_C, fill=Ecosystem)) + 
    geom_boxplot(aes(group=Ecosystem, value=Allo_C)) + #add proportion bars with black outline
    scale_x_discrete(guide = guide_axis(angle = 45)) +
    scale_fill_manual(values = wes_palette("Zissou1")) + #call from wesanderson library
    theme_few() #remove grid, white background, box around plot - from ggthemes

```


## References


Nave, L., K. Johnson, C. van Ingen, D. Agarwal, M. Humphrey, and N. Beekwilder. 2022. International Soil Carbon Network version 3 Database (ISCN3) ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/cc751923c5576b95a6d6a227d5afe8ba

